Modeling Two Body Collisions with Spring Potential Interactions

### 2014-12-13 paper/paper.tex
slight paper changes

Any two body collision can be reduced to 2-dimensional motion in a plane containing both masses and their initial velocity vectors. Additionally, in this paper we limit ourselves to the case where the initial velocities are parallel, though the masses may be initially off center from each other by a distance $s$, the impact parameter. One of the masses, the target $m_{1}$, may be at rest initially, though we also allow it to have an initial closing velocity in some studies. For continence we pick our coordinate system such that all the initial velocities lie along $\hat{x}$ and the target is on the $\hat{y}$ axis, $y_{1}(t=0)=0$ (Figure~\ref{fig:schematic}).

@@

The final velocity and maximum displacement results when varying the projectiles initial velocity and the inpact parameter largely matched our expectations. However, while varying the force power law, $\lambda$, the maximum displacement decreased with increasing $\lambda$ which was unexpected. We believe that this inverse relationship beween $\lambda$ and $\Delta x_{max}$ is due to all of our simulations having $\Delta x \leq 1$, additional studies could modify the initial parameters to investigate this theory further in the case where $\Delta x \geq 1$. When we included a non-zero energy disipation coefficent, $\eta$ we observed a very rapid decay of the total energy and hence final velocity. This general behvior was expected, but not to the degree seen. An improved method of including energy disipation for future simulations would be to normalize the energy lost per time step to $\Delta t$ by altering $\left(1-\eta\right)~\rightarrow~\left(1-\eta\right)\Delta t$ so that the energy loss per unit time is independent of the number of time steps used to simulate that unit time.

>>>>

Any two body collision can be reduced to 2-dimensional motion in a plane containing both masses and their initial velocity vectors. Additionally, in this paper we limit ourselves to the case where the initial velocities are parallel, though the masses may be initially off center from each other by a distance $s$, the impact parameter. One of the masses, the target $m_{1}$, may be at rest initially, though we also allow it to have an initial closing velocity in some studies. For convenience we pick our coordinate system such that all the initial velocities lie along $\hat{x}$ and the target is on the $\hat{y}$ axis, $y_{1}(t=0)=0$ (Figure~\ref{fig:schematic}).

@@

The final velocity and maximum displacement results when varying the projectiles initial velocity and the inpact parameter largely matched our expectations. However, while varying the force power law, $\lambda$, the maximum displacement decreased with increasing $\lambda$ which was unexpected. We believe that this inverse relationship beween $\lambda$ and $\Delta x_{\text{max}}$ is due to all of our simulations having $\Delta x \leq 1$, additional studies could modify the initial parameters to investigate this theory further in the case where $\Delta x \geq 1$. When we included a non-zero energy disipation coefficent, $\eta$ we observed a very rapid decay of the total energy and hence final velocity. This general behvior was expected, but not to the degree seen. An improved method of including energy disipation for future simulations would be to normalize the energy lost per time step to $\Delta t$ by altering $\left(1-\eta\right)~\rightarrow~\left(1-\eta\right)\Delta t$ so that the energy loss per unit time is independent of the number of time steps used to simulate that unit time.
